Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-04-14 manuscript/manuscript.tex
updated ms

@@

@@

@@

>>>>

We conducted extensive simulations to understand how temporal covariance,
$G(t)$, and convergence correlations behave under (1) different quantitative
genetic fitness models, (2) different trait architectures (e.g. varying levels
of $V_A$ for fitness and the number of sites affecting fitness) and (3)
background selection. Furthermore, we use two population simulations to
investigate how convergence correlations depend on (1) the population sizes of
each selection line sampled from the main population, and (2) the direction the
trait is selected on in each line (i.e. in the same direction, differing
directions, or only one lines elected).

Due to the high computational burden of forward simulations, we modeled a
single 50 megabase region in a population of $N = 1000$ diploid individuals
with a neutral variant mutation rate of $10^8$. This is roughly analogous to a
XXX chromosome of \emph{Drosophila melanogaster}; however with this small
population size and mutation rate, the population mutation rate $\theta$ for
the entire region leads to far fewer neutral sites to calculate covariances and
other statistics on than expected for a region this length in \emph{D.
melanogaster}. Since our main goal is to understand how the dynamics of
statistics used in the paper and how they are affected by background
selection, fitness model, and trait architecture, we use population
frequencies rather than sampling frequencies.

All forward simulations were conducted using SLiM \parencite{Haller2019-vu},
and the simulation routines are available in the Github repository
\url{https://github.com/vsbuffalo/cvtk/}.

@@

We first investigated the effects of the selected trait's genetic architecture
on temporal covariances and $G(t)$ by neutrally burning in a population for
$10N$ generations, and selecting on the trait with an exponential fitness
function. By using an exponential fitness function (a form of selection
multiplicative across sites), we ensure there is no build up of LD due to
selection (XXX) and epistasis; this serves as the simplest directional
selection model of a trait to understand the effects of genetic architecture on
the statistics we have used in the paper.

During this burnin, sites were either marked as neutral (with mutation rate
$\mu_\mathrm{neutral} = 10^8$ per gamete per generation) or contributed to the
trait's value (with mutation rate $\mu_\mathrm{trait}$), but were not selected
on until generation $10N + 5$ (the five generations after burnin serve as a
neutral control). The trait mutation rate, $\mu_\mathrm{trait}$ was set by
targeting a particular architecture, the number of selected sites, $L$. Each
site contributing to the trait's value was randomly chosen to have effect size
$\pm \alpha$ with equal probability, where $\alpha$ was set to target a
particular additive genetic variance for the trait, $V_A$, for the target
number of selected sites $L$.

Overall, we confirm a finding in \parencite{Buffalo2019-io} that the initial
expected temporal covariance conditioned on $V_A$, is invariant to the number
of loci determining the trait's value, $L$ (Supplementary Figure
\ref{suppfig:sim-expfit-covs}). We do find some evidence that the decay in
temporal covariance is faster when the trait has a monogenic basis (see the
third column of Supplementary Figure \ref{suppfig:sim-expfit-covs}); this is
expected the selection coefficients are larger for these monogenic simulations,
leading to faster allele frequency changes and a rapid change in additive
genetic variance.

In our previous work, we did not investigate the affect of trait architecture
on our measure $G(t)$.  Using the exponential fitness function simulations, we
also calculated $G(t)$ for each of the replicate simulations. We find that
across replicates, the $G(t)$ trajectories can vary considerably depending on
the number of sites ($L$) determining the trait's value (Supplementary Figure
\ref{suppfig:sim-expfit-Gs}). When a trait is monogenic ($L \approx 1$), $G(t)$
trajectories vary considerably across replicate lines, as certain lines may not
contain the few selected alleles (top row of Supplementary Figure
\ref{suppfig:sim-expfit-Gs}). However, with a polygenic trait, ($L \ge 100$),
the $G(t)$ trajectories across replicates are all quite close as each replicate
contains an abundance of alleles that affect the trait's value (bottom rows of
Supplementary Figure \ref{suppfig:sim-expfit-Gs}). Comparing the simulated
$G(t)$ replicate trajectories of Supplementary Figure
\ref{suppfig:sim-expfit-Gs} with the \textcite{Barghi2019-qy} $G(t)$
trajectories in Figure \ref{fig:figure-1}B, we again confirm a finding of
\textcite{Barghi2019-qy}: that there is considerable genetic redundancy among
beneficial alleles. We should note that our simplified simulation routines are
slightly different from the \textcite{Barghi2019-qy} study in that the burnin
populations are all independent; however we expect the same qualitative result.

\vb{TODO: note about different burnin pops?}

@@

  \caption{The temporal covariances $\cov(\Delta p_5, \Delta p_t)$ from the
    onset of selection (generation 5) to a later time point $t$, which varies
    along the x-axis, across a variety of different trait additive genetic
    variances ($V_A$, columns) and number of sites contributing to the trait
    ($L$, rows). Each point is the temporal covariance averaged over 50
    replicate simulations; dark gray points are temporal covariances after the
    onset of selection, and light gray points are before. The red line is a
    loess smoothed curve through the covariances after the onset of selection.
    Selection on the trait was imposed through an exponential fitness function.}

  \label{suppfig:sim-expfit-covs}
\end{figure}

@@

  \caption{The $G(t)$ trajectories of 50 replicate simulations, across
    different trait architectures ($L$ is the target number of sites affecting
    the trait's value, and $V_A$ is the target trait additive genetic
    variance).  The red line is the mean trajectory across all replicate
  simulations. Like Supplementary Figure \ref{suppfig:sim-expfit-covs}, the
onset of selection is five generations after the $10N$ generation burnin; this
is evident by the initial flat section of the $G(t)$ trajectory.}

  \label{suppfig:sim-expfit-Gs}
\end{figure}
